---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(leaps)
library(car)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
# Import the dataset
obesity_data <- read.csv("Obesity Classification.csv")

# Identify the character columns
char_cols <- sapply(obesity_data, is.character)

# Convert character columns to factors for easier analysis
obesity_data[char_cols] <- lapply(obesity_data[char_cols], as.factor)

```

* * *

## Part 1: Data

### How are the observations in the sample collected?
The provided dataset encompasses details regarding the classification of individuals' obesity. It has been compiled from diverse sources, including medical records, surveys, and self-reported data.
Due to this being an observational study, this analysis can only be used for generalized purposes; no causal connections can be made.

* * *

## Part 2: Research question
What metrics are most associated with weight?

* * *

## Part 3: Exploratory data analysis

#### Variables being considered
Dependent Variable
- BMI

Independent Variables
- Age
- Gender 
- Height 
- Label
- Weight

### Investigate the Dataset and Variables

```{r eda}
### Investigate the Dataset
head(obesity_data)
tail(obesity_data)
str(obesity_data)
summary(obesity_data)

### Investigate the variables
## Dependent Variable
summary(obesity_data$BMI)
barplot(table(obesity_data$BMI))
boxplot(table(obesity_data$BMI))
hist(table(obesity_data$BMI))

## Independent Variables
# Age
summary(obesity_data$Age)
hist(obesity_data$Age)
boxplot(obesity_data$Age)
scatter.smooth(obesity_data$Age)
qqnorm(obesity_data$Age)
qqline(obesity_data$Age)

# Gender
summary(obesity_data$Gender)
barplot(table(obesity_data$Gender))
boxplot(table(obesity_data$Gender))
hist(table(obesity_data$Gender))

# Height
summary(obesity_data$Height)
hist(obesity_data$Height)
boxplot(obesity_data$Height)
scatter.smooth(obesity_data$Height)
qqnorm(obesity_data$Height)
qqline(obesity_data$Height)

# Label
summary(obesity_data$Label)
barplot(table(obesity_data$Label))
boxplot(table(obesity_data$Label))
hist(table(obesity_data$Label))

# BMI
summary(obesity_data$Weight)
hist(obesity_data$Weight)
boxplot(obesity_data$Weight)
scatter.smooth(obesity_data$Weight)
qqnorm(obesity_data$Weight)
qqline(obesity_data$Weight)
```

### Data Manipulation and Cleaning
```{r manip_and_clean}
# Keep columns for analysis
obesity_data_trimmed <- obesity_data[, c("Weight","Age","Gender", "Height","BMI")]

# Check rows with N/A's and remove if feasible
num_rows_with_na <- sum(is.na(obesity_data_trimmed))

# Remove rows with N/As.
obesity_data_cleaned <- na.omit(obesity_data_trimmed)

# Analyze cleaned data
str(obesity_data_cleaned)
```

* * *

## Part 4: Modeling - Multiple Linear Regression

### Backwards Elimination - P-Value

```{r modeling}
# Full model
obesity_data_model_full = lm(BMI ~ Age + Gender + Height + Weight, data = obesity_data_cleaned)
summary(obesity_data_model_full)

# Model - Exclude Age
obesity_data_model_excludeOne = lm(BMI ~ Height + Gender + Weight, data = obesity_data_cleaned)
summary(obesity_data_model_excludeOne)

# Model - Exclude Height/Age
obesity_data_model_excludeTwo = lm(BMI ~ Gender + Weight, data = obesity_data_cleaned)
summary(obesity_data_model_excludeTwo)

# Final Model
obesity_data_model_final = lm(BMI ~ Gender + Weight, data = obesity_data_cleaned)
summary(obesity_data_model_final)
```

Height was left out of this model due to a low p-value score during backwards elimination. 
However, many models that calculate BMI to within a specific interval utilize height as a metric
for prediction. It could be that since this dataset is relatively small, the p-value did not appear
to be a statistically significant predictor. A much larger dataset is likely required to achieve
a better, more accurate model, that would likely include the Height metric as a predictor.

### Assess Collinearity of Predictor Variables in the Final Model
The adjusted GVIF taking into account the DF indicates a relatively low indication
of collinearity between predictor (independent) variables.
The GVID and adjusted GVIF are relatively low, and fall below the acceptable
threshold of 5. 

Another method would have been to use a correlation matrix, however, it only works with numerical
predictor variables.

```{r}
# Calculate VIF for the predictors in data_model_final
gvif_final <- vif(obesity_data_model_final)

# Calculate VIF for the predictors in data_model_excludeOne
gvif_excludeOne <- vif(obesity_data_model_excludeOne)

# Print the VIF - data_model_final
print(gvif_final)

# Print the VIF - data_model_excludeOne
print(gvif_excludeOne)
```

### Conditions

#### (1) linear relationships between (numerical) x and y
Check if each explanatory variable is linearly related to the response variable.

‣ Each (numerical) explanatory variable linearly related to the response variable
‣ Check using residuals plots (e vs. x)
‣ Looking for a random scatter around 0
‣ Instead of scatterplot of y vs. x: allows for considering the other variables 
that are also in the model, and not just the bivariate relationship between a 
given x and y.

```{r conditions_linear_relationships}
plot(obesity_data_model_final$residuals ~ obesity_data$Gender)
plot(obesity_data_model_final$residuals ~ obesity_data$Weight)
```

#### (2) nearly normal residuals with mean 0
The plots below indicate a relatively, nearly normal distribution of residuals scattered
around zero within the model.

‣ Some residuals will be positive and some negative
‣ On a residuals plot we look for random scatter of residuals around 0
‣ This translates to a nearly normal distribution of residuals centered at 0
‣ Check using histogram or normal probability plot

```{r}
hist(obesity_data_model_final$residuals)
qqnorm(obesity_data_model_final$residuals)
qqline(obesity_data_model_final$residuals)
```

#### (3) constant variability of residuals
Constant variance is effectively met, with a slight convergence indicated nearing
the right side of the residuals plot.

‣ Residuals should be equally variable for low and high values of the predicted 
response variable
‣ Check using residuals plots of residuals vs. predicted (e vs. )
‣ Residuals vs. predicted instead of residuals vs. x because it allows for 
considering the entire model (with all explanatory variables) at once
‣ Residuals randomly scattered in a band with a constant width around 0 (no 
fan shape)
‣ Also worthwhile to view absolute value of residuals vs. predicted to identify 
unusual observations easily

```{r}
plot(obesity_data_model_final$residuals ~ obesity_data_model_final$fitted)
plot(abs(obesity_data_model_final$residuals) ~ obesity_data_model_final$fitted)
```

#### (4) independent residuals
It does not appear there is time series structure involved with this data. The
data is a random sample of movies from IMBD and Rotten Tomatoes so independence should
be satisfied.

‣ Independent residuals → independent observations
‣ If time series structure is suspected check using residuals vs. order of data 
collection
‣ If not, think about how the data are sampled

```{r}
plot(obesity_data_model_final$residuals)
```

* * *

## Part 5: Prediction

```{r prediction}
# Calculate confidence intervals for the coefficients
conf_intervals_excludeOne <- confint(obesity_data_model_excludeOne)
conf_intervals_final <- confint(obesity_data_model_final)

# View the confidence intervals
conf_intervals_final
conf_intervals_excludeOne

## Models

# Final Model + Height
obesity_data_model_excludeOne = lm(BMI ~ Gender + Weight + Height, data = obesity_data_cleaned)

## Enter data for the model
# Test the model against observed values - Final Model + Height Predictor
test_one_excludeOne = data.frame(Gender = c("Male"), Weight = c(80), Height = c(175))
test_two_excludeOne = data.frame(Gender = c("Female"), Weight = c(60), Height = c(160))
test_three_excludeOne = data.frame(Gender = c("Male"), Weight = c(90), Height = c(180))
test_four_excludeOne = data.frame(Gender = c("Female"), Weight = c(50), Height = c(150))
test_five_excludeOne = data.frame(Gender = c("Male"), Weight = c(100), Height = c(190))

## Make predictions based on the movie data above - Final Model + Height Predictor
test_one_pred_excludeOne <- predict(obesity_data_model_excludeOne, test_one_excludeOne)
test_two_pred_excludeOne <- predict(obesity_data_model_excludeOne, test_two_excludeOne)
test_three_pred_excludeOne <- predict(obesity_data_model_excludeOne, test_three_excludeOne)
test_four_pred_excludeOne <- predict(obesity_data_model_excludeOne, test_four_excludeOne)
test_five_pred_excludeOne <- predict(obesity_data_model_excludeOne, test_five_excludeOne)

## View the predicted values - Final Model + Height Predictor
test_one_pred_excludeOne
test_two_pred_excludeOne
test_three_pred_excludeOne
test_four_pred_excludeOne
test_five_pred_excludeOne

# Final Model
obesity_data_model_final = lm(BMI ~ Gender + Weight, data = obesity_data_cleaned)

# Test the model against observed values - Final Model
test_one_final = data.frame(Gender = c("Male"), Weight = c(80))
test_two_final = data.frame(Gender = c("Female"), Weight = c(60))
test_three_final = data.frame(Gender = c("Male"), Weight = c(90))
test_four_final = data.frame(Gender = c("Female"), Weight = c(50))

## Make predictions based on the movie data above - Final Model
test_one_pred_final <- predict(obesity_data_model_final, test_one_final)
test_two_pred_final <- predict(obesity_data_model_final, test_two_final)
test_three_pred_final <- predict(obesity_data_model_final, test_three_final)
test_four_pred_final <- predict(obesity_data_model_final, test_four_final)

## View the predicted values - Final Model
test_one_pred_final
test_two_pred_final
test_three_pred_final
test_four_pred_final
```

* * *

## Part 6: Conclusion

When comparing the model results to some of the observed data, it appears to be fairly close. The confidence conf_intervals
for the predictors calculated earlier also seem to correspond to the variability in the predictions.
To improve the model, Logistic Regression can be utilized and a different response variable can be used, such as Label,
which indicates whether an individual is Underweight, Normal, Overweight or Obese.
A much larger dataset should also be used to build a better model which, as stated above, would likely end up including
the Height predictor variable, which appeared to not be statistically significant.